Kepler's Demographic Cliff: Ages of Young Stars and Planets in the Kepler Field

### 2024-05-01 paper/ms.tex
update ms

\author{and coauthors!}
%
%% Early experiments for Kepler rotation samples
%\author[0000-0001-7967-1795]{Elsa~K.~Palumbo}
%\affiliation{\caltech}
%
%% ?
%%\author{Lynne~A.~Hillenbrand}
%%\affiliation{\caltech}
%
%% HIRES observations and reductions
%\author[0000-0002-0531-1073]{H. Isaacson}
%\affiliation{\berkeley}
%% HIRES observations, reductions, JUMP infrastructure
%\author[0000-0001-8638-0320]{A. W. Howard}
%\affiliation{\caltech}
%%
%\author{HIRES observers to be named}
%\affiliation{Various}

% <250 words
\begin{abstract}
  Recent studies of the rotation rates and lithium content of FGK
  stars in open clusters have revealed how these quantities evolve
  during the first few billion years of a star's life.
  %
  Here we apply this knowledge to stars observed by Kepler in order to
  derive rotation and lithium-based ages.
  %
  Rotation periods are drawn from previous Kepler analyses; lithium is
  measured from new and archival HIRES spectra.
  %
  We report rotation-based ages for \nuniqstarsantosrotgyroappl\ stars
  and \nplwgyroagenograzing\ planets for which we believe
  gyrochronology is valid;
  these rotation-based ages are $\approx$15-20$\times$ more precise
  than isochrone ages for our sample of FGK stars.
  %
  The independent lithium and rotation-based ages agree for
  $\approx$90\% of stars.  However, the timescales of lithium
  depletion are such that rotation is the more precise age diagnostic
  for the majority of our stellar sample.
  %
  Our resulting catalog of planetary ages includes
  \nplyounggyrotwosigma\ planets younger than 1\,Gyr at 2$\sigma$
  (\nplyounggyrothreesigma\ planets at 3$\sigma$).
  %
  More broadly, we find that the age distribution of stars in the
  Kepler field decreases linearly from 3\,Gyr ago to present day,
  for both known planet hosts as well as the parent stellar sample.
  %
  While this stellar age distribution has implications for the
  star formation history of our Galaxy, it also bounds the relative
  balance of young and old planets detectable using both Kepler and
  also future wide-field transit surveys.
  %
\end{abstract}

@@

Given that Kepler's primary mission ended in 2013, one might be
forgiven for assuming that the ages of the Kepler planets are already
known.  Of course, they have been studied:
\citeauthor{Berger_2020b_rpage} and \citeauthor{Petigura_2022} for
instance calculated isochronal ages for individual stars, which is
most precise for stars whose luminosities and temperatures separate
them from the main-sequence.  \citeauthor{David_2021} leveraged both
isochronal ages from an earlier analogous study
\citep{Fulton_Petigura_2018_cks_vii}, and, in their core analysis,
this work used a relative gyrochrone age-ranking approach that sorted
stars based on whether they rotated faster or slower than a gyrochrone
age of $\approx$2.6\,Gyr \citep{Meibom_2015,Curtis_2020}.  These
studies were conducted in the broader context of investigating trends
in exoplanet demographics, which have been shown to depend on both
stellar metallicity and mass \citet{Petigura_2018}, at levels likely
stronger than age (at least over gigayear timescales).  However, these
studies offered limited sensitivity to the ages of the youngest
planets, due to the limitations of isochronal dating.

@@

{\it Candidate pulsators and close-in binaries (bit
10)}---\citeauthor{Santos_2021} included in a flag for candidate
``classical pulsators'' (e.g. RR~Lyrae, Cepheids) and candidate
close-in binaries.  Visual inspection of the rotation periods as a
function of temperature shows that although this flag selects many
bona fide objects in these classes, it also selects the overwhelming
majority of $\lesssim$5\,day bona fide young planets.  While we
propagate this flag into our set of quality flags, we therefore
generally ignore it.

{\it How do we use the bit-flags?)}
An important decision for our subsequent analysis is how to combine
the above quality indicators in order to decide whether or not a star is suited for
gyrochronology.  By default, we assume that a star is suitable for
gyrochronology if none of bits zero through nine (inclusive) are
raised.

\subsection{Planet Quality Flags}
\label{subsec:plflags}
For the subset of our analysis that focuses on the properties of
planets, we use the following additional quality indicators to assess 
the quality of a planet, $Q_{\rm planet}$.

% see Erik's 2022 discussion; might want to shift to include a cut on
% disposition score.
{\it Candidate reliability (bit 0)}---no false positives; only
``planet candidates'' and ``confirmed planets''.

{\it Candidate S/N (bit 1)}---the false positive rate increases
greatly toward the noise floor for planet detection in any transit survey \citep[e.g.][]{2002ApJ...564..495J}.  We
required a S/N in excess of Kepler's usual 7.1$\sigma$ floor, through a cut on the
``multiple event statistic'' (MES): we required ${\rm MES}$$>$10.

{\it Grazing planets (bit 2)}---grazing objects ($b$>0.8) often
provide non-robust planetary sizes due to the size-impact parameter
degeneracy.  Particularly for large planets, they can also include
astrophysical false positives at higher rates (due to the same
degeneracy).  While we include a flag for such objects, we by default
do not exclude them from our analysis; this is because...{\bf
justify}.

@@

The rotation periods for the stellar and planetary samples are shown
in Figure~\ref{fig:prot_vs_teff}.  Out of the \nkeplerstars\ stars
observed by Kepler, requiring reported rotation periods and a reliable
DR3 crossmatch yielded \nuniqstarsantosrot\ stars.  Further requiring
these rotating stars to have temperatures 3800$<$$T_{\rm
eff}$/K$<$6200, which is the range over which \texttt{gyro-interp}
returns finite age posteriors, yields \nuniqstarsantosrotteffcut\
stars.  Further requiring them to be apparently single, near the main
sequence, with $\log g$$>$4.2, $M_{\rm G}>3.9$, and not flagged as
eclipsing binaries yielded \nuniqstarsantosallbutruwe\ rotators
potentially amenable for gyrochronology.  Finally, imposing RUWE$<$1.4
and requiring no possible blends ($\Delta G$<2.5) within 4$''$ leaves
\nuniqstarsantosrotgyroappl\ stars for which gyrochronology is likely
to be valid.

In the planet sample, the cumulative KOI table included
\nnonfpcumkois\ confirmed and candidate planets, orbiting
\nnonfpcumkoihosts\ stars.  Assume that we apply all the same stellar
cuts, save for that on RUWE.  Then, require these confirmed and
candidate planets to have MES$>$10.  This combination of selection
criteria yield \nplwgyroagewithgrazingandhighruwe\ planets with
rotation-based ages orbiting
\nplhoststarwgyroagewithgrazingandhighruwe\ stars
(\nplhoststarwgyroagejusthighruwe\ of these stars with ``high'' RUWE).
If we additionally require the planets to be non-grazing ($b<0.8$)
based on the Markov Chain Monte Carlo (MCMC) fits from
\citet{Thompson_2018}, we are left with \nplwgyroagenograzing\ planets and
\nplhoststarwgyroagenograzing\ stars, respectively.

@@

Figure~\ref{fig:li_vs_teff} shows our measured lithium equivalent
widths, plotted over the mean isochrone models from
\texttt{EAGLES} \citep{Jeffries_2023}.  We declare an upper-limit for
plotting purposes if ${EW}-\sigma_{\rm EW}$$<$10\,m\AA.  However, our
EWs span -40 to 250\,m\AA; all negative EWs have uncertainties that
are statistically consistent with zero.

To calculate lithium ages, we use \texttt{EAGLES} (git commit
\texttt{ac09637}), which uses a similar method as
\texttt{gyro-interp}, and was calibrated on stars observed by the
Gaia-ESO spectroscopic survey in 52 open clusters with ages spanning 2
to 6{,}000\,Myr \citep{Jeffries_2023}.  We adopt a linear prior in
age, spanning $10^6$ to $10^{10}$\,yr, and symmetrize our EW
uncertainties for purposes of interfacing with \texttt{EAGLES} by
taking the maximum of the measured positive and negative 1$\sigma$
uncertainty intervals.
Certain quality flags described in Section~\ref{subsec:flags} are also
relevant for lithium ages, specifically those associated with
crowding and binarity.
For instance, rapid rotators are more common in binaries, and they also
generally have higher lithium abundances than slow-rotating stars of the
same mass (CITE).
Stars for which rotation-based age dating is dubious heavily overlap
with those for which lithium-based ages are dubious.

@@

\begin{figure*}[!t]
  \begin{center}
    \leavevmode
    \includegraphics[width=\textwidth]{hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240430_maxage3200.pdf}
  \end{center}
  \vspace{-0.6cm}
  \caption{
    {\bf Kepler's demographic cliff,} visible in the rotation-derived
    ages of stars (left) and planet hosts (right) in the Kepler field.
    The lightly colored histograms correspond to all
    \nuniqstarsantosrotteffcut\ $T_{\rm eff}\in[3800,6200]\,$K Kepler
    stars with reported rotation periods, and all
    \nplhoststarwgyroage\ planet-hosting stars with reported rotation
    periods.
    The darker histograms impose necessary quality cuts on factors
    such as binarity, crowding, and the star's evolutionary state
    (Section~\ref{subsec:flags}).
    In detail, these ``histograms'' are resampled from the posterior
    probability distributions.
    Completeness is near unity until $\approx$3\,Gyr
    \citep{2022ApJ...937...94M}.
    {\bf todo: numbers dont match.  why?}
    \label{fig:hist_tgyro}
  }
\end{figure*}

@@

\startlongtable
\begin{deluxetable*}{llllllllrrrrrrl}
  %\tabletypesize{\footnotesize}
  \tabletypesize{\scriptsize}
  \tablecaption{Ages of Kepler planets and planet candidates.  This
  version of the table is truncated to include the youngest fifty
  confirmed Kepler planets, and the twenty youngest candidates,
  sorted by the minimum of either the rotation or lithium-based
  age.  The
  full machine-readable table contains ages and age limits for
  \nnonfopkoissomeageinfo\ non-false positive KOIs with MES$>$10.  
  \label{tab:planets}}
  %\toprule
  %\midrule
  %\endhead
  \startdata
  KOI & Kepler &  $T_{\rm eff}$ & $P_{\rm rot}$ & EW$_{\rm Li}$ &
  $t_{\rm gyro}$ & $t_{\rm Li} $ & Consistent? & $R_{\rm p}$ & $P_{\rm orb}$ &
  $Q_{\rm planet}$ & $Q_{\rm gyro}$ &  Spec? & Comment \\
  -- &   -- & K & days & m\AA & Myr & Myr &  str &   Earths &    days &       int  & int & bool & -- \\
  \hline
  \input{table_subgyr_confirmed} %\\
% \rule
  \input{table_subgyr_candidate}
  \enddata
  \tablecomments{``Confirmed'' planets appear before``candidate'' planets.
    The bit quality flags for the rotation-based ages,
    $Q_{\rm gyro}$, are described in Section~\ref{subsec:flags}.
    Concisely summarized, they are:
    Bit 0: $T_{\rm eff}/{\rm K} \in [ 3800-6200]$ ?
    Bit 1: $\log g$$<$4.2?
    Bit 2: $M_{\rm G}$<3.9?
    Bit 3: In KEBC?
    Bit 4: Large $d_{\rm xm,Kep-Gaia}$?
    Bit 5: Confused Kep-Gaia crossmatch?
    Bit 6: Gaia DR3 non-single star?
    Bit 7: RUWE$>$1.4?
    Bit 8: Crowded?
    Bit 9: Far from main-sequence?
    Bit 10: \citetalias{Santos_2021} CP/CB?
    Bit 11: $P_{\rm rot}$ not in the homogeneous
    \citetalias{Santos_2019} \citetalias{Santos_2021} sample?
    As an example, Kepler-1627 has $Q_{\rm gyro}$ flagged with bit 10 and bit 7.
    The analogous planet quality bit-flag, $Q_{\rm planet}$, has the
    following meaning.
    Bit 0: Candidate not reliable?
    Bit 1: MES$<$10?
    Bit 2: Grazing?
    For a star to have a high likelihood of being ``reliable for gyrochronology'' we suggest $Q_{\rm star}$ bits
    zero to nine to not be raised, and for a planet to be ``reliabe'', we suggested $Q_{\rm planet}$ to be zero.
  }
\end{deluxetable*}

@@

For Kepler-957, $T_{\rm eff}$=4947\,K, and $P_{\rm rot}$=4.52\,days,
This implies a rotation-based age of $96_{-72}^{+192}$\,Myr.  The
rotational broadening of the spectrum ($v\sin
i$$\approx$9.8$\pm$1.0\kms) is consistent with the nominal stellar
rotation period.  However, if this star were $\lesssim$500\,Myr old,
lithium should have been detectable at $>$50\,m\AA.  Lithium is
marginally if at all present in the spectrum ({\bf  EW$_{\rm
Li}$$<$XX\,m\AA at 2$\sigma$ }).  While the planet (nominally
6.9\,$R_\oplus$ on a 5.9\,day orbit) has been statistically validated
\citep{Morton_2016}, inspecting the data validation reports this is
somewhat surprising.  The light curve shows greater scatter in-transit
than out of transit, appears near-grazing, and has a suggestion of a
strong odd-even depth difference.  This is a surprising system that
merits further attention.

>>>>

\author{coauthors!}

%
%% Early experiments for Kepler rotation samples
%\author[0000-0001-7967-1795]{Elsa~K.~Palumbo}
%\affiliation{\caltech}
%
%% ?
%%\author{Lynne~A.~Hillenbrand}
%%\affiliation{\caltech}
%
%% HIRES observations and reductions
%\author[0000-0002-0531-1073]{H. Isaacson}
%\affiliation{\berkeley}
%% HIRES observations, reductions, JUMP infrastructure
%\author[0000-0001-8638-0320]{A. W. Howard}
%\affiliation{\caltech}
%%
%\author{HIRES observers to be named}
%\affiliation{Various}

% <250 words
\begin{abstract}
  Recent studies of FGK stars in open clusters have quantitatively
  shown how stellar rotation rates and lithium content evolve over
  the first few billion years.
  %
  Here we apply this knowledge to stars observed by Kepler to compute
  rotation and lithium-based ages.
  %
  Rotation periods are drawn from previous Kepler analyses; lithium is
  measured from new and archival HIRES spectra.
  %
  We report rotation-based ages for \nuniqstarsantosrotgyroappl\ stars
  and \nplwgyroagenograzing\ planets for which we believe
  gyrochronology is valid;
  these rotation-based ages are $\approx$15-20$\times$ more precise
  than isochrone ages in our sample of FGK stars.
  %
  The independent lithium and rotation-based ages agree for
  $\approx$90\% of stars.  However, the timescales of lithium
  depletion are such that rotation is the more precise age diagnostic
  for the majority of our stellar sample.
  %
  Our resulting catalog of planetary ages includes
  \nplyounggyrotwosigma\ planets younger than 1\,Gyr at 2$\sigma$
  (\nplyounggyrothreesigma\ planets at 3$\sigma$).
  %
  More broadly, we find that the age distribution of stars in the
  Kepler field decreases linearly from 3\,Gyr ago to present day,
  for both planet hosts as well as for the overall stellar sample.
  %
  While this stellar age distribution has implications for the
  star formation history of our Galaxy, it also bounds the relative
  balance of young and old planets detectable through
  transit surveys of nearby stars in the thin disk.
  %
\end{abstract}

@@

Given that Kepler's primary mission ended in 2013, one might be
forgiven for assuming that the ages of the Kepler planets are already
known.  Of course, they have been studied:
\citeauthor{Berger_2020b_rpage} and \citeauthor{Petigura_2022} for
instance calculated isochronal ages for individual stars, which is
most precise for stars whose luminosities and temperatures separate
them from the main-sequence.  \citeauthor{David_2021} leveraged both
isochronal ages from an earlier analogous study
\citep{Fulton_Petigura_2018_cks_vii}, and, in their core analysis,
this work used a relative gyrochrone age-ranking approach that sorted
stars based on whether they rotated faster or slower than a gyrochrone
age of $\approx$2.6\,Gyr \citep{Meibom_2015,Curtis_2020}.  These
studies were conducted in the broader context of investigating trends
in exoplanet demographics, which have been shown to depend on both
stellar metallicity and mass \citep{Petigura_2018}, at levels likely
stronger than age (at least over gigayear timescales).  However, these
studies offered limited sensitivity to the ages of the youngest
planets, due to the limitations of isochronal dating.

@@

{\it Candidate pulsators and close-in binaries (bit
10)}---\citeauthor{Santos_2021} included in a flag for candidate
``classical pulsators'' (e.g.\ RR~Lyrae, Cepheids) and candidate
close-in binaries.  Visual inspection of the rotation periods as a
function of temperature shows that although this flag selects many
bona fide objects in these classes, it also selects the overwhelming
majority of $\lesssim$5\,day bona fide young planets.  In our view this is
erroneous because they these objects are neither classical pulsators nor
close binaries.  While we
propagate this flag into our set of quality flags, we therefore
generally ignore it.

{\it How do we use the bit-flags}
An important decision for our subsequent analysis is how to combine
the above quality indicators in order to decide whether or not a star is suited for
gyrochronology.  By default, we assume that a star is suitable for
gyrochronology if none of bits zero through nine (inclusive) are
raised.

\subsection{Planet Quality Flags}
\label{subsec:plflags}
Not all planets are born equal.
We use the following additional quality indicators to assess 
the reliability and utility of a planet, and assemble them into a separate bit-flag,
$Q_{\rm planet}$.

% see Erik's 2022 discussion; might want to shift to include a cut on
% disposition score.
{\it Candidate reliability (bit 0)}---The NEA's Kepler Objects of Interest Table
includes both an overall planet-candidate disposition status
({\it koi\_disposition}), as well as a disposition based only on the Kepler
data ({\it koi\_pdisposition}).
We require both to include only
``planet candidates'' and ``confirmed planets''; we are generally less interested
in the age distribution of planet candidates that are known false positives.

{\it Candidate S/N (bit 1)}---In any transit survey, the false positive rate increases
greatly toward the noise floor for planet detection \citep[e.g.][]{2002ApJ...564..495J}.  We
required a S/N in excess of Kepler's usual 7.1$\sigma$ floor, through a cut on the
maximum
``multiple event statistic'' (MES, {\it koi\_max\_mult\_ev}): we required ${\rm MES}$$>$10.

{\it Grazing planets (bit 2)}---Grazing objects, for which the impact
parameter $b$ is greater than $1-R_{\rm p}/R_\star$,
often yield biased planetary parameters
\citep[e.g.][]{2022AJ....163..111G}.
Particularly for large planets, they can also include
astrophysical false positives at higher rates \citep{2016ApJ...822...86M}, which is
in part due to the size-impact parameter degeneracy.
We flag planet candidates as potentially grazing if $b>0.8$, 
using the impact parameters reported in the Markov Chain
Monte Carlo (MCMC) fits from \citet{Thompson_2018}.

@@

The rotation periods for the stellar and planetary samples are shown
in Figure~\ref{fig:prot_vs_teff}.  Out of the \nkeplerstars\ stars
observed by Kepler, requiring reported rotation periods and a reliable
DR3 crossmatch yielded \nuniqstarsantosrot\ stars.  Further requiring
these rotating stars to have temperatures 3800$<$$T_{\rm
eff}$/K$<$6200, which is the range over which \texttt{gyro-interp}
returns finite age posteriors, yields \nuniqstarsantosrotteffcut\
stars.  Further requiring the stars to be apparently single, near the main
sequence, with $\log g$$>$4.2, $M_{\rm G}>3.9$, and not flagged as
eclipsing binaries yielded \nuniqstarsantosallbutruwe\ rotators
potentially amenable for gyrochronology.  Finally, imposing RUWE$<$1.4
and requiring no possible blends ($\Delta G$<2.5) within 4$''$ leaves
\nuniqstarsantosrotgyroappl\ stars for which gyrochronology is likely
to be valid.

In the planet sample, the cumulative KOI table included
\nnonfpcumkois\ confirmed and candidate planets, orbiting
\nnonfpcumkoihosts\ stars.  Assume that we apply all the same stellar
cuts, save for that on RUWE.  Then, require these confirmed and
candidate planets to have MES$>$10.  This combination of selection
criteria yield \nplwgyroagewithgrazingandhighruwe\ planets with
rotation-based ages orbiting
\nplhoststarwgyroagewithgrazingandhighruwe\ stars
(\nplhoststarwgyroagejusthighruwe\ of these stars with ``high'' RUWE).
If we additionally require the planets to be non-grazing ($b<0.8$)
based on the Markov Chain Monte Carlo (MCMC) fits from
\citet{Thompson_2018}, we are left with \nplwgyroagenograzing\ planets orbiting
\nplhoststarwgyroagenograzing\ stars.

@@

Figure~\ref{fig:li_vs_teff} shows our measured lithium equivalent
widths, plotted over the mean isochrone models from
\texttt{EAGLES} \citep{Jeffries_2023}.  We declare an upper-limit for
plotting purposes if the 1$\sigma$ lower limit on the equivalent width
is below 10\,m\AA.  Our measured
EWs span -40 to 250\,m\AA; all negative EWs have uncertainties that
are statistically consistent with zero.

To calculate lithium ages, we use \texttt{EAGLES} (git commit
\texttt{ac09637}), which, similar to
\texttt{gyro-interp}, is based on a purely empirical interpolation
approach.  \texttt{EAGLES} however was calibrated on lithium measurements of
stars observed by the
Gaia-ESO spectroscopic survey in 52 open clusters with ages spanning 2
to 6{,}000\,Myr \citep{Jeffries_2023}.  We adopt a linear prior in
age, spanning $10^6$ to $10^{10}$\,yr, and symmetrize our EW
uncertainties for purposes of interfacing with \texttt{EAGLES} by
taking the maximum of the measured positive and negative 1$\sigma$
uncertainty intervals.
Certain quality flags described in Section~\ref{subsec:flags} are also
relevant for lithium ages, specifically those associated with
crowding and binarity.
For instance, rapid rotators are more common in binaries, and they also
generally have higher lithium abundances than slow-rotating stars of the
same mass (CITE).
Stars for which rotation-based age dating is dubious heavily overlap
with those for which lithium-based ages are dubious.

@@

\begin{figure*}[!t]
  \begin{center}
    \leavevmode
    \includegraphics[width=\textwidth]{hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240430_maxage3200.pdf}
  \end{center}
  \vspace{-0.6cm}
  \caption{
    {\bf Kepler's demographic cliff,} visible in the rotation-derived
    ages of stars (left) and planet hosts (right) in the Kepler field.
    The lightly colored histograms correspond to all
    \nuniqstarsantosrotteffcut\ $T_{\rm eff}\in[3800,6200]\,$K Kepler
    stars with reported rotation periods, and all
    \nplhoststarwgyroage\ planet-hosting stars with reported rotation
    periods.
    The darker histograms impose necessary quality cuts on factors
    such as binarity, crowding, and the star's evolutionary state
    (Section~\ref{subsec:flags}).
    In detail, these ``histograms'' are resampled from the posterior
    probability distributions.
    Completeness is near unity at t$\lesssim$3\,Gyr
    \citep{2022ApJ...937...94M}.
    {\bf todo: numbers dont match.  why?}
    \label{fig:hist_tgyro}
  }
\end{figure*}

@@

\startlongtable
\begin{deluxetable*}{llllllllrrrrrrl}
  %\tabletypesize{\footnotesize}
  \tabletypesize{\scriptsize}
  \tablecaption{Ages of Kepler planets and planet candidates.  This
  version of the table is truncated to include only the youngest confirmed Kepler planets,
  sorted by the minimum of either the rotation or lithium-based
  age.  The
  full machine-readable table contains ages and age limits for
  \nnonfopkoissomeageinfo\ non-false positive KOIs with MES$>$10.  
  A shell script to decode the $Q_{\rm gyro}$ quality flag is {\bf \href{https://gist.github.com/lgbouma/20368253f1a98da1b39cf32fdda0be13}{available online}}.
  \label{tab:planets}}
  %\toprule
  %\midrule
  %\endhead
  \startdata
  KOI & Kepler &  $T_{\rm eff}$ & $P_{\rm rot}$ & EW$_{\rm Li}$ &
  $t_{\rm gyro}$ & $t_{\rm Li} $ & Consistent? & $R_{\rm p}$ & $P_{\rm orb}$ &
  $Q_{\rm planet}$ & $Q_{\rm gyro}$ &  Spec? & Comment \\
  -- &   -- & K & days & m\AA & Myr & Myr &  str &   Earths &    days &       int  & int & bool & -- \\
  \hline
  \input{table_subgyr_confirmed} %\\
% \rule
%  \input{table_subgyr_candidate}
  \enddata
  \tablecomments{``Confirmed'' planets appear in the machine-readable version before``candidate'' planets.
    The bit quality flags for the rotation-based ages,
    $Q_{\rm gyro}$, are described in Section~\ref{subsec:flags}.
    Concisely summarized, they are:
    Bit 0: $T_{\rm eff}/{\rm K} \in [ 3800-6200]$ ?
    Bit 1: $\log g$$<$4.2?
    Bit 2: $M_{\rm G}$<3.9 or $M_{\rm G}>8.5$?
    Bit 3: In KEBC?
    Bit 4: Large $d_{\rm xm,Kep-Gaia}$?
    Bit 5: Confused Kep-Gaia crossmatch?
    Bit 6: Gaia DR3 non-single star?
    Bit 7: RUWE$>$1.4?
    Bit 8: Crowded?
    Bit 9: Far from main-sequence?
    Bit 10: \citetalias{Santos_2021} CP/CB?
    Bit 11: $P_{\rm rot}$ not in the homogeneous
    \citetalias{Santos_2019} \citetalias{Santos_2021} sample?
    As an example, Kepler-1627 has $Q_{\rm gyro}$ flagged with bit 10 and bit 7.
    The analogous planet quality bit-flag, $Q_{\rm planet}$, has the
    following meaning.
    Bit 0: Candidate not reliable?
    Bit 1: MES$<$10?
    Bit 2: Grazing?
    For a star to have a high likelihood of being ``reliable for gyrochronology'' we suggest $Q_{\rm star}$ bits
    zero to nine to not be raised, and for a planet to be ``reliabe'', we suggested $Q_{\rm planet}$ to be zero.
  }
\end{deluxetable*}

@@

For Kepler-957, $T_{\rm eff}$=4947\,K, and $P_{\rm rot}$=4.52\,days,
This implies a rotation-based age of $96_{-72}^{+192}$\,Myr.  The
rotational broadening of the spectrum ($v\sin
i$$\approx$9.8$\pm$1.0\kms) is consistent with the nominal stellar
rotation period.  However, if this star were $\lesssim$500\,Myr old,
lithium should have been detectable at $>$50\,m\AA.  Lithium is
marginally if at all present in the spectrum ({\bf  EW$_{\rm
Li}$$<$XX\,m\AA at 2$\sigma$ }).  While the planet (nominally
6.9\,$R_\oplus$ on a 5.9\,day orbit) has been statistically validated
\citep{2016ApJ...822...86M}, inspecting the data validation reports this is
somewhat surprising.  The light curve shows greater scatter in-transit
than out of transit, appears near-grazing, and has a suggestion of a
strong odd-even depth difference.  This is a surprising system that
merits further attention.
